---
title: "Human Activity Recognition Using Machine Learning"
author: "Lutvi Nurhadiansyah"
date: "2/13/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **1. Load Required Libraries**

```{r load-libraries}
# Install necessary packages if not already installed
if (!require(caret)) install.packages("caret", dependencies = TRUE)
if (!require(randomForest)) install.packages("randomForest", dependencies = TRUE)
if (!require(ggplot2)) install.packages("ggplot2", dependencies = TRUE)
if (!require(dplyr)) install.packages("dplyr", dependencies = TRUE)

# Load libraries
library(caret)
library(randomForest)
library(ggplot2)
library(dplyr)
```

## **2. Load the Data**

```{r load-data}
# Load the dataset
train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

train_data <- read.csv(url(train_url), na.strings = c("NA", "", "#DIV/0!"))
test_data <- read.csv(url(test_url), na.strings = c("NA", "", "#DIV/0!"))

# View the structure of the dataset
str(train_data)
```

## **3. Data Cleaning**

```{r data-cleaning}
# Remove columns with excessive NA values
train_data <- train_data[, colSums(is.na(train_data)) == 0]

# Remove irrelevant columns (e.g., user_name, timestamps, row numbers)
train_data <- train_data[, -c(1:7)]
test_data <- test_data[, colnames(test_data) %in% colnames(train_data)]

# Convert 'classe' to a factor (target variable)
train_data$classe <- as.factor(train_data$classe)

# Check cleaned data
str(train_data)
```

## **4. Data Splitting**

```{r data-splitting}
set.seed(123)
trainIndex <- createDataPartition(train_data$classe, p = 0.7, list = FALSE)
training <- train_data[trainIndex, ]
validation <- train_data[-trainIndex, ]

# Check class distribution
table(training$classe)
```

## **5. Train a Model (Random Forest)**

```{r model-training}
set.seed(123)
rf_model <- randomForest(classe ~ ., data = training, importance = TRUE, ntree = 100)

# View model summary
print(rf_model)
```

## **6. Model Evaluation**

```{r model-evaluation}
# Predict on validation set
predictions <- predict(rf_model, validation)

# Compute accuracy
conf_matrix <- confusionMatrix(predictions, validation$classe)
print(conf_matrix)
```

## **7. Predict on Test Data**

```{r test-predictions}
# Make predictions on test set
test_predictions <- predict(rf_model, test_data)

# View predictions
print(test_predictions)
```

## **8. Conclusion**
This project successfully applies a **Random Forest Classifier** to predict human activity based on sensor data. The model achieves high accuracy and can be improved further by **hyperparameter tuning** or using advanced techniques like **XGBoost**.
